### 2014-03-02 paper/concrete.tex
Comments on section 4, small changes

The operational semantics we have defined in the previous section
satisfy non-interference by design.
We achieve this general statement that works for a large class of
languages by having different task execute completely isolated from
each other, such that every task has its own state.
In some cases, this strong separation is desirable, or even necessary.
Languages like C provide direct access to memory locations without
mechanisms in the language to achieve a separation of the heap.
On the other hand, for other languages this
strong isolation of tasks can be
undesirable, e.g., for performance reasons.
For instance, for the language |targetLangML|, our presentation so far
requires a separate heap per task, which is not very practical.
Instead, we would like to
more tightly couple the integration of the target and IFC
language by reusing existing infrastructure.  In the running example,
a concrete implementation might use a single global heap.
More precisely, instead of using a configuration of the form
\[|iconf iS (fullconf id1 il1 tS1 ie1, fullconf id2 il2 tS2 ie2 ldots)|\]
we would like a single global heap as in
\[|oneheapiconf iS tS (oneheapfullconf id1 il1 ie1, oneheapfullconf id2 il2 ie2, ldots)|\]

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated.  In particular, it is not
sound to share references to a heap cell between different tasks.
Otherwise, information can be trivially leaked by copying
secret data into a reference in one task, and then
reading it again by a another task with a lower label.
What we would like is a way of characterizing safe modifications to
the semantics which preserve non-interference.
The intention of our single heap implementation was to be more efficient
by using one global heap, but conceptually maintain isolation between
task by not allowing sharing of references between tasks.
This intuition of having a different (potentially more efficient)
concrete semantics that behave like the abstract semantics
can be formalized by the following definition:

@@

A na\"ive attempt at our single heap semantics
will not allow us to construct an isomorphism
with the abstract language |specLangML alpha|, and instead highlights
where things will go wrong.
To successfully construct an isomorphism, our concrete semantics need
to mirror the behavior of the abstract semantics, most importantly
in the case where tasks interact.  Namely, at a fork and when sending
or receiving messages, we take an expression from one task and
execute it in the context of another task.  In the abstract semantics
the fork rule is parameterized by the strategy |klone|, which defines
what heap the new task should execute in.  \Red{In Section~\ref{sec:retrofit}
we have motivated two choices}, namely the identity (to effectively
clone the heap), or to start with an empty heap.
In the latter case, this effectively means that existing addresses
in the forked expression |ie| are no longer valid and the
task gets stuck (and then removed by \textsc{I-noStep}).
\todo{}{Is this actually true?  Currently the semantics are unclear.}
To get the isomorphic behavior in the concrete semantics with a
single heap, we would have to change the forked expression |ie| in
such a way that all addresses get rewritten to fresh addresses
that are guaranteed to not occur in the global heap.  Similarly,
sending a memory address should take a similar approach, as addresses
for sent references will not exist in the receiving task.

@@

Again we are modifying the IFC language semantics in a non-trivial way,
which begs the question whether non-interference is preserved.
Note that restricting the transitions in an information flow control
system in general does not preserve non-interference as the following
minimal example illustrates.
Consider a trivial state machine as shown
in Figure~\ref{fig:trivial-sm}, whose second projection is classified
secret.  With all three transitions, this language fulfills
termination sensitive non-interference.  Furthermore, if the
dashed transition
is removed, the language continues to satisfy TSNI.  However, if any
solid line is removed, the language fails TSNI.

@@

\begin{definition}[Restricted IFC language]
  \label{def:restricted}
  For a family of predicates $\mathcal P$ (one for every reduction
  rule), we call
  |restrictedLang alpha targetLang| a restricted IFC language
  if its definition is equivalent to the abstract language
  |specLang alpha targetLang|, with the following exception:
  The reduction rules are restricted
  by adding a predicate $P$ from $\mathcal P$ to the premise of
  all rule other than \textsc{I-noStep}.  Furthermore, the predicates $P$
  can depend only on the \textit{erased} configuration
  |erase il ic|, where |il| is the label of the first task
  in the task list and |ic| the full configuration.
\end{definition}

@@

\begin{theorem}
  \label{thm:restricted}
  For any target language |targetLang| and family of predicates
  $\mathcal{P}$, the IFC language |restrictedLang roundrobinf targetLang|
  is TSNI.  Furthermore, the IFC language
  |restrictedLang seqf targetLang| is TSNI.
\end{theorem}

@@

We are now ready to make our single heap IFC language precise and
ensure its non-interference using the techniques presented.
First, we can construct the restricted language
|restrictedLangNoRef alpha targetLangML|, where |noRefs| is
the family of always valid predicates, except for the ones for
\textsc{I-fork} and \textsc{I-send}, which we define as
\[ P = \text{|ie| does not contain any address |ta|} \]
That is, we do not restrict any rules except for \textsc{I-fork}
and \textsc{I-send}.
Since $P$ only depends on |ie|, which is part of the current
task and thus never erased w.r.t.\ the label of the first task,
this language satisfies non-interference by Theorem~\ref{thm:restricted}.

>>>>

The operational semantics we have defined in the previous section
satisfy non-interference by design.
We achieve this general statement that works for a large class of
languages by having different tasks executing completely isolated from
each other, such that every task has its own state.
In some cases, this strong separation is desirable, or even necessary.
Languages like C provide direct access to memory locations without
mechanisms in the language to achieve a separation of the heap.
On the other hand, for other languages this
strong isolation of tasks can be
undesirable, e.g., for performance reasons.
For instance, for the language |targetLangML|, our presentation so far
requires a separate heap per task, which is not very practical.
Instead, we would like to
more tightly couple the integration of the target and IFC
language by reusing existing infrastructure.  In the running example,
a concrete implementation might use a single global heap.
More precisely, instead of using a configuration of the form
\[|iconf iS (fullconf id1 il1 tS1 ie1, fullconf id2 il2 tS2 ie2 ldots)|\]
we would like a single global heap as in
\[|oneheapiconf iS tS (oneheapfullconf id1 il1 ie1, oneheapfullconf id2 il2 ie2, ldots)|\]

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated.  In particular, it is not
sound to share references to a heap cell between different tasks.
Otherwise, information can be trivially leaked by copying
secret data into a reference in one task, and then
reading it again by a another task with a lower label.
What we would like is a way of characterizing safe modifications to
the semantics which preserve non-interference.
The intention of our single heap implementation was to be more efficient
by using one global heap, but conceptually maintain isolation between
task by not allowing sharing of references between them.
This intuition of having a different (potentially more efficient)
concrete semantics that behave like the abstract semantics
can be formalized by the following definition:

@@

\Red{the next paragraph needs some rephrasing. We should say that if we 
condense everything into one environment, then we need to ``simulate'' 
the different policies. A bit more intro}
A na\"ive attempt at our single heap semantics
will not allow us to construct an isomorphism
with the abstract language |specLangML alpha|, and instead highlights
where things will go wrong.
To successfully construct an isomorphism, our concrete semantics need
to mirror the behavior of the abstract semantics, most importantly
in the case where tasks interact.  Namely, at a fork and when sending
or receiving messages, we take an expression from one task and
execute it in the context of another task.  In the abstract semantics
the fork rule is parameterized by the strategy |klone|, which defines
what heap the new task should execute in.  \Red{In Section~\ref{sec:retrofit}
we have motivated two choices}, namely the identity (to effectively
clone the heap), or to start with an empty heap.
In the latter case, this effectively means that existing addresses
in the forked expression |ie| are no longer valid and the
task gets stuck (and then removed by \textsc{I-noStep}).
\todo{}{Is this actually true?  Currently the semantics are unclear.}
To get the isomorphic behavior in the concrete semantics with a
single heap, we would have to change the forked expression |ie| in
such a way that all addresses get rewritten to fresh addresses
that are guaranteed to not occur in the global heap.  Similarly,
sending a memory address should take a similar approach, as addresses
for sent references will not exist in the receiving task.

@@

Again, we are modifying the IFC language semantics in a non-trivial way,
which begs the question whether non-interference is preserved.
Restricting the transitions in an information flow control
system in general does not preserve non-interference as the following
minimal example illustrates. \Red{Ale: I am not sure that the example is clear,
  too abstract, at least, I don't get it. Propose to remove it, and just say 
the top level idea: ``the restriction should not be based on secret data!''} 
Consider a trivial state machine as shown
in Figure~\ref{fig:trivial-sm}, whose second projection is classified
secret.  With all three transitions, this language fulfills
termination sensitive non-interference.  Furthermore, if the
dashed transition
is removed, the language continues to satisfy TSNI.  However, if any
solid line is removed, the language fails TSNI.

@@

\begin{definition}[Restricted IFC language]
  \label{def:restricted}
  For a family of predicates $\mathcal P$ (one for every reduction
  rule), we call
  |restrictedLang alpha targetLang| a restricted IFC language
  if its definition is equivalent to the abstract language
  |specLang alpha targetLang|, with the following exception:
  The reduction rules are restricted
  by adding a predicate $P \in \mathcal P$ to the premise of
  all rule other than \textsc{I-noStep}.  Furthermore, the predicates $P$
  can depend only on the \textit{erased} configuration
  |erase il ic|, where |il| is the label of the first task
  in the task list and |ic| the full configuration.
\end{definition}

@@

\begin{theorem}
  \label{thm:restricted}
  For any target language |targetLang| and family of predicates
  $\mathcal{P}$, the restricted IFC language |restrictedLang roundrobinf targetLang|
  is TSNI.  Furthermore, the IFC language
  |restrictedLang seqf targetLang| is TINI.
\end{theorem}

@@

We are now ready to make our single heap IFC language precise and
ensure its non-interference using the techniques presented.
First, we can construct the restricted language
|restrictedLangNoRef alpha targetLangML|, where |noRefs| is
the family of always valid predicates, except for the ones for
\textsc{I-fork} and \textsc{I-send}, which we define as
\[ P(e) = \text{|ie| does not contain any address |ta|} \]
That is, we do not restrict any rules except for \textsc{I-fork}
and \textsc{I-send}.
Since $P$ only depends on |ie|, which is part of the current
task and thus never erased w.r.t.\ the label of the first task,
this language satisfies non-interference by Theorem~\ref{thm:restricted}.
